\chapter{Setting Up}
 \index{Setting Up}
       

To play Dungeon World, you'll need to gather yourself and 2–5 friends. A group of 4 to 5, including you, is best. Choose one person to be the GameMaster (hereafter: GM). Everyone else will be players, taking the role of the characters in the game. The players get to say what their characters say, think, and do. The GM describes everything else in the world.

       

You can play a single session or string together multiple sessions into a campaign. Plan accordingly. Each session will usually be a few hours and you'll be able to start playing right away within the first session.

       

You'll need to print some materials. Before you start a new game, print off at least:

       
\startitemize[1,packed]
         
\item A few copies of the basic and special moves

         
\item One copy of each class sheet

         
\item One copy each of the cleric and wizard spell sheets

         
\item One copy of the adventure sheet and GM moves

       
\stopitemize
       

Everyone at the table will need something to write with and some six-sided dice. Two dice is the minimum but two dice per player is a good idea.

       

You'll also need some specialized dice: four-sided, eight-sided, and ten-sided. One of each is enough but more is better; you won't have to pass them around so much.

       
\section{Using Dice}  \index{Using Dice} \index{Dice}
       

When talking about dice, {\bf dX}  means a dice with X sides (d4, d6, d8, etc.). A number before the dice size means "roll this many and add them together," so 2d6 means "roll two six-sided dice and add them together." Sometimes you'll add something to the roll, maybe a static number or the result of another roll, add that to the total. If there's a "·b" at the end instead of adding them together you take the best one ('b' for 'best'), so 2d6·b+1 means "roll two six-sided dice and take the best one and add one to it." A "·w" at the end means you take the worst result, so 2d6·w means "roll two six-sided dice and take the worst one."

       
