\chapter{Cleric Spells}
 \index{Cleric Spells} \index{Cleric} \index{Spells}
 


\section{Rotes}  \index{Rotes} \index{Rotes}
 


\startSpellName
\CMSpellName{Light} 	Rote
\stopSpellName
 

An item you touch glows with divine light, about as bright as a torch. It gives off no heat or sound and requires no fuel but is otherwise like a mundane torch. You have complete control of the color of the flame. The spell lasts as long as it is in your presence.



 
\startSpellName
\CMSpellName{Sanctify} 	Rote
\stopSpellName
 

Food or water you hold in your hands while you cast this spell is consecrated by your deity. In addition to now being holy or unholy the affected substance is purified of any mundane spoilage.

 
\startSpellName
\CMSpellName{Guidance} 	Rote
\stopSpellName
 

The symbol of your deity appears before you and gestures towards the direction or course of action your deity would have you take then disappears. The message is through gesture only; your communication through this spell is severely limited.



 


\section{1st Level Spells}  \index{1st Level Spells} \index{1st} \index{Level} \index{Spells}
 
\startSpellName
\CMSpellName{Bless} 	Level 1	{\em Ongoing} 
\stopSpellName
 

You deity smiles on the target in combat. They take +1 ongoing so long as battle continues and they stand and fight. While this spell is ongoing you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Cure Light Wounds} 	Level 1
\stopSpellName
 

At your touch wounds scab and bones cease to ache. Heal an ally of 1d8 damage.

 
\startSpellName
\CMSpellName{Detect Alignment} 	Level 1
\stopSpellName
 

When you cast this spell choose an alignment: Good, Evil, or Neutral. One of your senses is briefly able to detect that alignment. The GM will tell you what here is of that alignment.

 
\startSpellName
\CMSpellName{Cause Fear} 	Level 1	{\em Ongoing} 
\stopSpellName
 

Choose an intelligent target you can see and a nearby object. The target is afraid of the object so long as you maintain the spell. Their reaction is up to them: flee, panic, beg, panic, fight. While this spell is ongoing you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Magic Weapon} 	Level 1	 Ongoing
\stopSpellName
 

The weapon you hold while casting does +1d4 damage until you dismiss this spell. Until you dismiss this spell you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Sanctuary} 	Level 1
\stopSpellName
 

You make an area holy to your deity. Walk the perimeter of the area. So long as you stay within that area you are alerted whenever someone acts with malice within the sanctuary (including entering with harmful intent). Anyone who receives healing within a Sanctuary heals +1d4 HP.

 
\startSpellName
\CMSpellName{Speak With Dead} 	Level 1	{\em Death} 
\stopSpellName
 

A corpse converses with you briefly. It will answer any three questions you pose to it to the best of the knowledge it had in life and the knowledge it gained in death.



 


\section{3rd Level Spells}  \index{3rd Level Spells} \index{3rd} \index{Level} \index{Spells}
 
\startSpellName
\CMSpellName{Animate Dead} 	Level 3	{\em Ongoing} 
\stopSpellName
 

You invoke a hungry spirit to possess a recently-dead body and act for you. This forms a zombie that follows your orders to the best of its limited abilities. Treat the zombie as your character, but with access to only the basic moves. It has a +1 modifier for all stats and 1 HP. While this spell is ongoing you take -1 to Cast a Spell. You get 1d4 of these effects:

 
\startitemize[1,packed]

\item The zombie is talented. Give one stat a +2 modifier.

 
\item The zombie is durable. It has +2 HP for each level you have.

 
\item The zombie has a functioning brain and can complete complex tasks.

 
\item The zombie is restored by magic—it does not appear obviously dead, at least for a day or two.


\stopitemize
 
\startSpellName
\CMSpellName{Cure Moderate Wounds} 	Level 3
\stopSpellName
 

You staunch bleeding and set bones through magic. Heal an ally of 2d8 damage.

 
\startSpellName
\CMSpellName{Darkness} 	Level 3	{\em Ongoing} 
\stopSpellName
 

Choose an area you can see: it's filled with supernatural darkness and shadow. While this spell is ongoing you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Resurrection} 	Level 3
\stopSpellName
 

Tell the GM you would like to resurrect a corpse whose soul has not yet fully departed this world. The GM will tell you "yes, you can resurrect them, but first…" or "yes, you can resurrect them now, but it won't be permanent until…" and then one to all of the things from this list:

 
\startitemize[1,packed]

\item It's going to take days/weeks/months

 
\item You must \_\_\_\_

 
\item You must get help from \_\_\_\_

 
\item It will require a lot of money

 
\item You must sacrifice \_\_\_\_ to do it


\stopitemize
 
\startSpellName
\CMSpellName{Hold Person} 	Level 3
\stopSpellName
 

Choose a creature you can see. Until you Cast a Spell or leave their presence they cannot act except to speak. If they're harmed this effect ends.



 


\section{5th Level Spells}  \index{5th Level Spells} \index{5th} \index{Level} \index{Spells}
 
\startSpellName
\CMSpellName{Revelation} 	Level 5
\stopSpellName
 

Your deity answers your prayers with a moment of perfect understanding. The GM will explain the root cause of the current situation. When acting on the information, you take +1 Forward.

 
\startSpellName
\CMSpellName{Cure Critical Wounds} 	Level 5
\stopSpellName
 

Heal an ally of 3d8 damage.

 
\startSpellName
\CMSpellName{Divination} 	Level 5
\stopSpellName
 

Name a person, place, or thing you want to learn about. Your deity grants you visions of the target, as clear as if you were there.

 
\startSpellName
\CMSpellName{Contagion} 	Level 5	{\em Ongoing} 
\stopSpellName
 

While this spell is ongoing a target you can see suffers from the disease of your choice and you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Words of the Unspeaking} 	Level 5
\stopSpellName
 

With a touch you speak to the spirits within things. The non-living object you touch responds to three questions you pose, answering them as best it can.

 
\startSpellName
\CMSpellName{True Seeing} 	Level 5	{\em Ongoing} 
\stopSpellName
 

While this spell is ongoing your vision is opened to the true nature of everything you lay your eyes on and you take -1 to Cast a Spell. You pierce illusions and see things that have been hidden. The GM will describe the area before you ignoring any illusions and falsehoods, magical or otherwise.

 
\startSpellName
\CMSpellName{Trap Soul} 	Level 5
\stopSpellName
 

When cast in the presence of a ghost or recently dead body this spell traps the soul in a gem you provide. While trapped the soul answers every question posed to it and can not resist your requests. Once released the soul is likely to hold a grudge against its captor.

 




\section{7th Level Spells}  \index{7th Level Spells} \index{7th} \index{Level} \index{Spells}
 
\startSpellName
\CMSpellName{Word of Recall} 	Level 7
\stopSpellName
 

Choose a word. The first time after casting this spell that you speak the chosen word, you and any allies touching you when you cast the spell are immediately returned to the exact spot you cast the spell at. Casting Word of Recall again before speaking the word replaces the earlier recall.

 
\startSpellName
\CMSpellName{Heal} 	Level 7
\stopSpellName
 

Touch an ally and you may restore up to your maximum HP to them.

 
\startSpellName
\CMSpellName{Harm} 	Level 7
\stopSpellName
 

Touch an enemy and strike them with divine wrath—deal 2d8 damage to them and 1d6 damage to yourself (ignores armor).

 
\startSpellName
\CMSpellName{Sever} 	Level 7	{\em Ongoing} 
\stopSpellName
 

Choose an appendage on the target such as an arm, tentacle, or wing. The appendage is magically severed from their body, causing no damage but considerable pain. Missing an appendage may, for example, keep a winged creature from flying, or a bull from goring you on its horns. While you maintain the spell you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Mark of Death} 	Level 7
\stopSpellName
 

Choose a creature who's true name you know. This spell inscribes runes that will kill that creature, should they read them.

 
\startSpellName
\CMSpellName{Control Weather} 	Level 7
\stopSpellName
 

Pray for rain—or sun, wind, or snow. Within a day or so, your god will answer. The weather will change according to your will and last a handful of days.



 


\section{9th Level Spells}  \index{9th Level Spells} \index{9th} \index{Level} \index{Spells}
 
\startSpellName
\CMSpellName{Command Elements} 	Level 9
\stopSpellName
 

A mass of one element (earth, air, water, or fire) that you touch as you cast this spell takes on a form similar to yours and carries out one command you give it. Once the command is completed or a day passes the element retakes its old form.

 
\startSpellName
\CMSpellName{Repair} 	Level 9
\stopSpellName
 

Choose one event in the target's past. All effects of that event, including damage, poison, disease, and magical effects, are ended and repaired. HP and diseases are healed, poisons are neutralized, magical effects are ended.

 
\startSpellName
\CMSpellName{Divine Presence} 	Level 9	{\em Ongoing} 
\stopSpellName
 

Every creature must ask your leave to enter your presence, and you must speak permission for them to enter. Any creature that you deny permission takes an extra 1d10 damage whenever they take damage in your presence. While this spell is ongoing you take -1 to Cast a Spell.

 
\startSpellName
\CMSpellName{Consume Unlife} 	Level 9
\stopSpellName
 

The mindless undead creature you touch is destroyed and you steal its death energy to heal yourself or the next ally you touch of damage equal to its current HP.

 
\startSpellName
\CMSpellName{Peace} 	Level 9
\stopSpellName
 

Choose one traumatic memory in the target's past. The target's memory of that event is calmly erased. If the target is a PC, they must be willing.







 
