\chapter{Magic Items}
 \index{Magic Items} \index{Magic} \index{Items}
       

There are stranger things in the world than swords and leather. Magic items are the non-mundane, items that have intrinsic power.

       

Magic items are for you to make for your game. Players can make magic items through the Wizard's ritual and similar moves. The GM can introduce magic items in the spoils of battle or the bounty of reward. This list provides some ideas, but magic items are ultimately for you to decide.

       

When making your own magic items keep in mind that magic items are magical. +1 damage is the realm of the mundane, magic items should provide more interesting bonuses.

       

Bag of Holding	\CMTags{0 Weight} 

       

A bag of holding is larger on the inside than the outside, it can contain an infinite number of items, and its weight never increases. When you try to retrieve an item from a bag of holding, roll+Wis. On a 10+, it's right there. On a 7-9, choose one:

       
\startitemize[1,packed]
         
\item You get the exact item, but it takes a while

         
\item You get a similar item of the GM's choice, but it only takes a moment

       
\stopitemize
       

No matter how many items it contains, a bag of holding is always 0 weight.

       

The Burning Wheel	\CMTags{2 Weight} 

       

This wagon wheel with several shattered spokes is constantly wreathed in flames. A gift from the gods, it takes a strong soul to carry the wheel. It does not provide any bonus to swimming.

       

When you pick up The Burning Wheel and speak a god's name, roll+Con. On a 7+, the god you name takes notice and grants you an audience. An audience with a god is not without a price: on a 10+, you choose one of your stats and reduce it to the next lowest modifier (for example, a 14 is +1, so it would be reduced to 11, a +0). On a 7-9, the GM chooses which stat to reduce.

       

Candle of Truth	\CMTags{0 Weight} 

       

So long as this candle burns, no one can willingly lie within its light.

       

Common Scroll	\CMTags{1 Use, 0 Weight} 

       

A common scroll has a spell inscribed on it. The spell must be castable by you or on your class's spell list for you to be able to cast it. When you cast a spell from a scroll, the spell has effect, simple as that.

       

Cornucopia	\CMTags{1 Weight} 

       

This horn of plenty converts magic into sustenance. When you cast a spell into a cornucopia, it produces food enough for a number of people equal to the spell's level.

       

Crystal Ball	\CMTags{1 Weight} 

       

When you look into a crystal ball, name a location and roll+Wis. On a hit, you get a clear vision of the location so long as you stare into the crystal ball. On a 7-9, someone or something else also observes you through the crystal ball.

       

Demonsbane	\CMTags{1 Use, 0 Weight} 

       

A weapon bathed in Demonsbane causes great pain to demons. It's touch will drive them back.

       

Flask of Breath	\CMTags{0 Weight} 

       

This flask contains an everlasting supply of air. It is useful for breathing underwater or on the far planes.

       

Holy Avenger	\CMTags{Close, 2 Weight} 

       

This sword is revered among paladins. It is considered a sign of divine favor to wield one. When wielded by a paladin, the paladin is granted a mark of divinity and the eyes of the gods are always on them (for good or ill). For anyone else, it’s just a sword.

       

Immovable Rod	\CMTags{0 Weight} 

       

When you press the small button on the end of the immovable rod, it freezes in place and cannot be moved. When you press the button again, it becomes movable again.

       

Infinite Book	\CMTags{1 Weight} 

       

This book contains an infinite number of pages in a finite space. With no limit to the pages, everything that ever was, is, or will be is contained somewhere in the book. Luckily the index is great.

       

When you Spout Lore while consulting the book you gain an extra clause. On a 12+ the GM will give a solution to a problem or situation you're in.

       

Inspectacles	\CMTags{0 Weight} 

       

A pair of rough-hewn glass lenses in shabby frames, they somehow allow the wearer to see much more than is actually present. When you Discern Realities while wearing the glasses, you have an extra clause. On a roll of 12+ you may ask any three questions about the person, situation or area you are observing, not just those on the list.

       

Living Arrows	\CMTags{1 Ammo, 1 Weight} 

       

These arrows know how to find their target. When you Volley, you may choose to deplete your smart arrow ammo to automatically succeed before you roll.

       

Lodestone Shield	\CMTags{1 Weight} 

       

Emblazoned with a roaring lion, this shield has the power to attract blades and arrows to it. When you Defend with a Lodestone Shield, when you spend hold against an enemy wielding a metallic weapon you may spend one more hold to also disarm them.

       

Mental Wax	\CMTags{1 Use, 0 Weight} 

       

When you drip mental wax into someone's ear, you get 3 Hold. You may spend hold to ask the player of the target one question which they must answer truthfully.

       

Rope of Tricks	\CMTags{1 Weight} 

       

This rope responds to its wielder’s every wish, tying itself into knots, grabbing on to items (or enemies), coiling tightly. It acts as if it was an intelligent creature, like a smart snake of rope. Every rogue wants one.

       

Teleportation Room	\CMTags{Slow} 

       

It requires an entire stone room to contain these powerful teleportation magics. When you enter the teleportation room and name a location, roll. On a 10+, you arrive exactly where you intended. On a 7-9, the GM chooses a safe location nearby. On a 6-, you end up someplace nearby, but it surely isn't safe.

       

Timunn's Armor	\CMTags{Armor 2, 1 Weight} 

       

This suit of armor appears as many things to many people. Its wearer always looks impeccably dressed to everyone who sees them. The aura of the armor makes everyone more likely to see it the wearer's way.

       

When you Parley with someone they see you as who they want you to be, you have a new clause for this move. On a roll of 12+ the target becomes hopelessly devoted to you. Tell the GM to stop making moves with this character, from now on they will only act in your best interest. (GM, don’t mess with this. Really, best interests.)

       

Vellius’s Gauntlets	\CMTags{1 Weight} 

       

Bearing the insignia of Saint Vellius, these gloves shine with the power of their wearer. When you hold something with the gloves, you cannot be forced to drop what you hold through force.

       

Vorpal Sword	\CMTags{3 Piercing, 2 Weight} 

       

When you deal damage with the Vorpal Sword, your enemy must choose something (an item, an advantage, a limb) and lose it, permanently.

       

Wax Wings	\CMTags{1 Weight} 

       

Crafted to resemble great eagle wings, this contraption is worn via a complex harness. When you take to the air with wax wings, roll+Dex. On a 10+ your flight is controlled and you may stay aloft as long as you like. On a 7-9, you make it aloft but your flight is short or erratic and unpredictable, your choice. On a 6- you make it aloft, but the rest is up to the GM.

                
