\chapter{An Example Front: The Opening of the White Gate}
 \index{An Example Front The Opening of the White Gate} \index{Front} \index{White} \index{Gate}
 
\section{Dangers}  \index{Dangers} \index{Dangers}
 
\subsection{The College of Arcanists (Cabal)}  \index{The College of Arcanists (Cabal)} \index{College} \index{Arcanists} \index{(cabal)}
 

{\bf Impulse:}  to absorb those in power, to grow

 
\subsubsection{Grim Portents}  \index{Grim Portents} \index{Grim} \index{Portents}
 
\startitemize[1,packed]

\item The College sends an expedition to the Gate

 
\item The Key is discovered

 
\item The Gate's Power is harnessed

 
\item The College seizes control


\stopitemize
 

{\bf Impending Doom:}  Usurpation

 
\subsection{The White Gate (Dark Portal)}  \index{The White Gate (Dark Portal)} \index{White} \index{Gate} \index{(dark} \index{Portal)}
 

{\bf Impulse:}  to disgorge demons

 
\subsubsection{Grim Portents}  \index{Grim Portents} \index{Grim} \index{Portents}
 
\startitemize[1,packed]

\item The First Trumpet sounds

 
\item The Second Trumpet sounds

 
\item The Gate is opened


\stopitemize
 

{\bf Impending Doom:}  Destruction

 
\subsection{The Argent Seraphim (Choir of Angels)}  \index{The Argent Seraphim (Choir of Angels)} \index{Argent} \index{Seraphim} \index{(choir} \index{Angels)}
 

{\bf Impulse:}  to pass judgement

 
\subsubsection{Grim Portents}  \index{Grim Portents} \index{Grim} \index{Portents}
 
\startitemize[1,packed]

\item A Champion is chosen

 
\item An organization of power is formed or co-opted

 
\item The Herald appears

 
\item Judgement is passed


\stopitemize
 

{\bf Impending Doom:}  Tyranny

 
\section{Description and Cast}  \index{Description and Cast} \index{Description} \index{Cast}
 

An ancient gate, buried for aeons in the icy north. It opens into a realm of pure light, guarded by the Argent Seraphim. It was crafted only to be opened at Judgement Day, so that the Seraphim could come forth and purge the realm of men. Recently uncovered by the College of Arcanists, who do not yet understand its terrible power.

 
\startitemize[1,packed]

\item Oren Balserus, Arcanist Supreme

 
\item Hali'el, voice of the Seraphim

 
\item Drudge, a manservant


\stopitemize
 
\section{Custom Moves}  \index{Custom Moves} \index{Custom} \index{Moves}
 

When you stand in the presence of the Light From Beyond, roll+Wis. On a 10+ you are judged worthy, the Argent Seraphim will grant you a vision or boon. On a 7-9 you are under suspicion and see a vision of what dark fate might befall you if you do not correct your ways.1On a miss, thou art weighed in the balance and art found wanting.

 
\section{Stakes}  \index{Stakes} \index{Stakes}
 
\startitemize[1,packed]

\item Who will be the Champion?

 
\item How will Lux respond to the holy light?

 
\item Will the College be able to recruit Avon?


\stopitemize
 

 




