\chapter{The Paladin}
 \index{The Paladin} \index{Paladin}
 





Hell awaits. An eternity of torment in fire or ice or whatever best suits the sins of the damned throngs of Dungeon World. All that stands between the pits of that grim torture and salvation is you. Holy man, armored war machine, templar of the Good and the Light, right? The Cleric may say his prayers at night to the gods, dwelling in their heavens. The Fighter may wield his sharp sword in the name of “good” but you know. Only you.

 

Eyes, hands, and sweet killing blow of the gods, you are. Yours is the gift of righteousness and virtue. Of justice. Vision, too. A purity of intent that your companions do not have. So guide these fools, Paladin. Take up your holy cause and bring salvation to the wastrel world.

  

Vae victis, right?



 
\section{Names}  \index{Names} \index{Names}
 



Thaddeus, Augustine, Lux, Cassius, Hadrian, Lucia, Octavia, Regulus, Valeria, Sanguinus, Titanius



 
\section{Look}  \index{Look}
 



Choose one for each:

 

Kind Eyes, Fiery Eyes, or Glowing Eyes

 

Helmet, Styled Hair, or Bald

 

Worn Holy Symbol or Fancy Holy Symbol

 

Fit Body, Bulky Body, or Thin Body



 
\section{Stats}  \index{Stats} \index{Stats}
 



Assign these scores to your stats:

 

17 (+2), 15 (+1), 13 (+1), 11 (+0), 9 (+0), 8 (-1)

 

You start with 10+Constitution HP.



 

Your base damage is d10.

 
\section{Starting Moves}  \index{Starting Moves} \index{Moves}
 




\startInstructionsAfterHeader
You are human, so you get this move.
\stopInstructionsAfterHeader
 
\subsection{Human}  \index{Human} \index{Human}
 

When you pray, even for a moment, to your deity and ask "What here is evil?" the GM will tell you, honestly.



 


\startInstructions
You start with these moves:
\stopInstructions
 
\subsection{Lay on Hands (Cha)}  \index{Lay on Hands (Cha)} \index{Lay} \index{Hands} \index{(cha)}
 

When you touch someone, skin to skin, and pray for your deity to cure them, roll+Cha. On a 10+ you heal 1d8 damage or remove one disease. On a 7-9, they are healed, but the damage or disease is transferred to you.

 
\subsection{Armored}  \index{Armored} \index{Armored}
 

You ignore the Clumsy tag on armor you wear.

 
\subsection{I Am the Law}  \index{I Am the Law} \index{Law}
 

When you give an NPC an order based on your divine authority, roll+Cha. On a hit, they choose one:

 
\startitemize[1,packed]

\item They do it

 
\item They back away cautiously, then flee

 
\item They attack you


\stopitemize
 

On a 10+, you also take +1 forward against them. On a miss, they do as they please and you take -1 Forward against them.

 
\subsection{Quest}  \index{Quest} \index{Quest}
 

When you dedicate yourself to a mission through prayer and ritual cleansing state what you set out to do:

 
\startitemize[1,packed]

\item Slay \thinrules[2], a great blight on the land

 
\item Defend \thinrules[2] from the inequities that beset them

 
\item Discover the truth of \thinrules[2]


\stopitemize
 

Then choose up to two boons:

 
\startitemize[1,packed]

\item An unwavering sense of direction to \thinrules[2].

 
\item Invulnerability to \thinrules[2] (ex: edged weapons, fire, enchantment, etc.)

 
\item A mark of divine authority

 
\item Senses that pierce lies

 
\item A voice that transcends language

 
\item A freedom from hunger, thirst, and sleep


\stopitemize
 

The GM will then tell you what vow or vows is required of you to maintain your blessing:

 
\startitemize[1,packed]

\item Honor (forbidden: cowardly tactics and tricks)

 
\item Temperance (forbidden: gluttony in food, drink, and pleasure of the flesh)

 
\item Piety (required: observance of daily holy services)

 
\item Valor (forbidden: suffering an evil creature to live)

 
\item Truth (forbidden: lies)

 
\item Hospitality (required: comfort to those in need, no matter who they are)


\stopitemize




 
\section{Alignment}  \index{Alignment} \index{Alignment}
 
\startInstructionsAfterHeader
Choose an alignment:
\stopInstructionsAfterHeader
 


\subsection{Lawful}  \index{Lawful} \index{Lawful}
 

Deny mercy to a criminal or unbeliever

 
\subsection{Good}  \index{Good} \index{Good}
 

Endanger yourself to protect someone weaker than you



 
\section{Gear}  \index{Gear} \index{Gear}
 



Your Load is 8+Str. You start with dungeon rations (1 weight, 5 uses), scale armor (2 armor, 3 weight), and some mark of faith, describe it (0 Weight). Choose your weapon:

 
\startitemize[1,packed]

\item Halberd (Reach, +1 damage, two-handed, 2 weight)

 
\item Long sword (Close, +1 damage, 1 weight) and shield (+1 armor, 2 weight)


\stopitemize
 

Choose one:

 
\startitemize[1,packed]

\item Adventuring gear (1 weight)

 
\item dungeon rations (1 weight) and healing potion


\stopitemize


 
\section{Bonds}  \index{Bonds} \index{Bonds}
 



Fill in the name of one of your companions in at least one:

 

\thinrules[2]'s misguided behavior endangers their very soul!

 

\thinrules[2] has stood by me in battle and can be trusted completely.

 

I respect the beliefs of \thinrules[2] but hope they will someday see the true way.

 

\thinrules[2] is a brave soul, I have much to learn from them.



 
\section{Advanced Moves}  \index{Advanced Moves} \index{Advanced} \index{Moves}
 


\startInstructionsAfterHeader
When you gain a level from 2-5, choose from these moves.
\stopInstructionsAfterHeader
 
\subsection{Divine Favor}  \index{Divine Favor} \index{Divine} \index{Favor}
 

You gain the Commune and Cast a Spell Cleric move. Your level for the purposes of that move is 1 + the levels you've gained since you took Divine Favor.

 
\subsection{Bloody Aegis}  \index{Bloody Aegis} \index{Bloody} \index{Aegis}
 

When you take damage you can use your body to deflect the blow. If you do you take no damage but instead suffer a debility of your choice. If you already have all six debilities you can't use this move.

 
\subsection{Smite}  \index{Smite} \index{Smite}
 

While on a Quest you deal +1d4 damage.

 
\subsection{Exterminatus}  \index{Exterminatus} \index{Exterminatus}
 

When you speak aloud your promise to defeat an enemy, you deal +2d4 damage against that enemy and -4 damage against anyone else. This effect lasts until the enemy is defeated or you admit your failure and prove your worth.

 
\subsection{Charge!}  \index{Charge} \index{Charge}
 

When you lead the charge into combat, those you lead take +1 forward.

 
\subsection{Staunch Defender}  \index{Staunch Defender} \index{Staunch} \index{Defender}
 

When you Defend you always get +1 hold. Even on a failure you get 1 hold.

 
\subsection{Setup Strike}  \index{Setup Strike} \index{Setup} \index{Strike}
 

When you Hack and Slash, choose an ally. Their next attack against your target does +1d4 damage.

 
\subsection{Holy Protection}  \index{Holy Protection} \index{Holy} \index{Protection}
 

You get +1 armor while on a Quest.

 
\subsection{Voice of Authority}  \index{Voice of Authority} \index{Voice} \index{Authority}
 

Take +1 to Order Hirelings.

 
\subsection{Hospitaller}  \index{Hospitaller} \index{Hospitaller}
 

When you heal an ally, you heal +1d8 damage.

 
\startInstructions
When you gain a level from 6-10, choose from these moves or the level 2-5 moves.
\stopInstructions
 
\subsection{Evidence of Faith}  \index{Evidence of Faith} \index{Evidence} \index{Faith}
 

Requires: Divine Favor

 

When you see divine magic as it happens, you can ask the GM which deity granted the spell and its effects. Take +1 when acting on the answers.

 
\subsection{Holy Smite}  \index{Holy Smite} \index{Holy} \index{Smite}
 

Replaces: Smite

 

While on a Quest you deal +1d8 damage.

 
\subsection{Ever Onward}  \index{Ever Onward} \index{Onward}
 

Replaces: Charge!

 

When you lead the charge into combat, those you lead take +1 forward and +2 armor forward.

 
\subsection{Impervious Defender}  \index{Impervious Defender} \index{Impervious} \index{Defender}
 

Replaces: Staunch Defender

 

When you Defend you always get +1 hold. Even on a failure you get 1 hold. When you get a 12+ to Defend instead of getting hold the nearest attacking creature is stymied giving you a clear advantage, the GM will describe it.

 
\subsection{Tandem Strike}  \index{Tandem Strike} \index{Tandem} \index{Strike}
 

Replaces: Setup Strike

 

When you Hack and Slash, choose an ally. Their next attack against your target does +1d4 damage and they take +1 forward against them.

 
\subsection{Divine Protection}  \index{Divine Protection} \index{Divine} \index{Protection}
 

Replaces: Holy Protection

 

You get +2 armor while on a Quest.

 
\subsection{Divine Authority}  \index{Divine Authority} \index{Divine} \index{Authority}
 

Replaces: Voice of Authority

 

Take +1 to Order Hirelings. When you roll a 12+ to Order Hirelings the hireling transcends their moment of fear and doubt to perform at the peak of potential.

 
\subsection{Perfect Hospitaller}  \index{Perfect Hospitaller} \index{Perfect} \index{Hospitaller}
 

Replaces: Hospitaller

 

When you heal an ally, you heal +2d8 damage.

 
\subsection{Indomitable}  \index{Indomitable} \index{Indomitable}
 

When you suffer a debility (even through Bloody Aegis) take +1 forward against whatever caused it.

 
\subsection{Perfect Knight}  \index{Perfect Knight} \index{Perfect} \index{Knight}
 

When you Quest you choose three boons instead of two.









 
