\chapter{Equipment}
 \index{Equipment} \index{Equipment}
 



The musty tombs and forgotten treasure troves of the world are filled with useful items. The Fighters can find a sharp new sword or the Thief might stumble across a deadly poison. Most items are mundane; not magical or intrinsically unique in any way. Any item that is magical or one-of-a-kind is not mundane for the purposes of moves. The Fighter's signature weapon is never mundane.

 

Weapons don't kill monsters, people do. That's why weapons don't have damage listed. A weapon is useful primarily for its tags which describe what the weapon is useful for. A dagger is not useful because it does more or less damage than some other blade. It's useful because its small and easy to strike with at close distance. A dagger in the hands of a Wizard is not nearly so dangerous as one in the hands of a skilled Fighter.

 

The stats below are for typical items. There are, of course, variations. A dull long sword might be -1 damage instead while a masterwork dagger could be +1 damage. Consider the following to be stats for a typical weapon of that type—a specific weapon could have different tags to represent its features.

 
\section{Tag Glossary}  \index{Tag Glossary} \index{Tag} \index{Glossary}
 

{\em n Ammo} : It counts as n ammo for appropriate ranged weapons.

 

{\em Applied} : It's only useful when carefully applied to a person or something they eat or drink.

 

{\em n Armor} : It protects you from harm and absorbs damage. When you take damage, subtract your armor from the total. If you have more than one item with n-armor, only the highest value counts. 

 

{\em +n Armor} : It protects you and stacks with other armor. Add its value to your total armor.

 

{\em Awkward} : It's unwieldy and tough to use. 

 

{\em +Bonus} : It modifies your effectiveness in some particular situation. It might be "+1 forward to Spout Lore" or "-1 ongoing to Hack and Slash."

 

{\em Clumsy} : It's tough to move around with. -1 ongoing while using it.

 

{\em n coins} : How much it costs to buy, normally. If the cost includes "-Charisma" a little negotiation subtracts the haggler's Charisma score (not modifier) from the price.

 

{\em +n Damage} : It is particularly harmful to your enemies. When you deal damage, you add n to it.

 

{\em Dangerous} : It's easy to get in trouble with it. If you interact with it without proper precautions the GM may freely invoke the consequences of your foolish actions.

 

{\em Forceful} : It can knock someone back a pace, maybe even off their feet. 

 

{\em Ignores Armor} : Don't subtract armor from the damage taken.

 

{\em Messy} : It does damage in a particularly destructive way, ripping people and things apart.

 

{\em n Piercing} : It goes right through armor. When you deal damage with n piercing, you subtract n from the enemy's armor for that attack. 

 

{\em Precise} : It rewards careful strikes. You use Dex to Hack and Slash with this weapon, not Str.

 

{\em Ration} : It's edible, more or less.

 

{\em Reload} : After you attack with it, it takes more than a moment to reset for another attack. 

 

{\em Requires} : It's only useful to certain people. If you don't meet the requirements it works weakly, if at all.

 

{\em Slow} : It takes minutes or more to use.

 

{\em Thrown} : Throw it at someone to hurt them. If you Volley with this weapon, you can't choose to mark off ammo on a 7–9. Once you throw it, it's lost until you can recover it. 

 

{\em Touch} : It's used by touching it to the target's skin.

 

{\em Two-handed} : It takes two hands to use it effectively.

 

{\em Stun} : When you attack with it, it does stun damage instead of normal damage. 

 

{\em n Uses} : It can only be used n times. 

 

{\em n Weight} : Count the listed amount against your Load. Something with no listed weight isn't designed to be carried. 100 coins in standard denominations is 1 weight. The same value in gems or fine art may be lighter or heavier.

 

{\em Worn} : To use it, you have to be wearing it.

 
\subsection{Range Tags (From Closest to Furthest)}  \index{Range Tags (From Closest to Furthest)} \index{Range} \index{Tags} \index{(from} \index{Closest} \index{Furthest)}
 

{\em Hand} : It's useful for attacking something within your reach, no further.

 

{\em Close} : It's useful for attacking something at arm's reach plus a foot or two.

 

{\em Reach} : It's useful for attacking something that's several feet away—maybe as far as ten. 

 

{\em Near} : It's useful for attacking if you can see the whites of their eyes. 

 

{\em Far} : It's useful for attacking something in shouting distance.

 
\section{Weapons}  \index{Weapons} \index{Weapons}
 

Ragged Bow	\CMTags{Near, 15 coins, 2 Weight} 

 

Fine Bow	\CMTags{Near, Far, 60 coins, 2 Weight} 

 

Hunter’s Bow	\CMTags{Near, Far, 100 coins, 1 Weight} 

 

Crossbow	\CMTags{Near, +1 damage, Reload, 35 coins, 3 Weight} 

 

Bundle of Arrows	\CMTags{3 Ammo, 1 coin, 1 Weight} 

 

Elven Arrows	\CMTags{4 Ammo, 20 coins, 1 Weight} 

 

Club, Shillelagh	\CMTags{Close, 1 coins, 2 Weight} 

 

Staff	\CMTags{Close, Two-handed, 1 coins, 1 Weight} 

 

Dagger, Shiv, Knife	\CMTags{Hand, 2 coins, 1 Weight} 

 

Throwing Dagger	\CMTags{Thrown, Near, 1 coin, 0 Weight} 

 

Short Sword, Axe, Warhammer, Mace	\CMTags{Close, 8 coins, 1 Weight} 

 

Spear	\CMTags{Reach, Thrown, Near, 5 coins, 1 Weight} 

 

Long Sword, Battle Axe, Flail	\CMTags{Close, +1 damage, 15 coins, 2 Weight} 

 

Halberd	\CMTags{Reach, +1 damage, Two-handed, 9 coins, 2 Weight} 

 

Rapier	\CMTags{Close, Precise, 25 coins, 1 Weight} 

 

Dueling Rapier	\CMTags{Close, 1 piercing, Precise, 50 coins, 2 Weight} 

 
\section{Armor}  \index{Armor} \index{Armor}
 

Leather, Chainmail	\CMTags{1 Armor, Worn, 10 coins, 1 Weight} 

 

Scale Mail	\CMTags{2 Armor, Worn, Clumsy, 50 coins, 3 Weight} 

 

Plate	\CMTags{3 Armor, Worn, Clumsy, 350 coins, 4 Weight} 

 

Shield	\CMTags{+1 Armor, 15 coins, 2 Weight} 

 
\section{Dungeon Gear}  \index{Dungeon Gear} \index{Dungeon} \index{Gear}
 

Adventuring Gear	\CMTags{5 Uses, 20 coins, 1 Weight} 

 

Adventuring gear is a collection of useful mundane items such as chalk, poles, spikes, ropes, etc. When you rummage through your adventuring gear for some useful mundane item, you find what you need and mark off a Use.

 

Bandages	\CMTags{3 Uses, Slow, 5 coins, 0 Weight} 

 

When you have a few minutes to bandage someone else's wounds, heal them of 4 damage and expend a use.

 

Poultices and Herbs	\CMTags{2 Uses, Slow, 10 coins, 1 Weight} 

 

When you carefully treat someone's wounds with poultices and herbs, heal them of 7 damage and expend a use.

 

Healing Potion	\CMTags{50 coins} 

 

When you drink an entire healing potion, heal yourself of 10 damage or remove one debility, your choice.

 

Keg of Dwarven Stout	\CMTags{10 coins, 4 Weight} 

 

When you open a keg of dwarven stout and let everyone drink freely, take +1 to your Carouse roll. If you drink a whole keg yourself, you are very very drunk.

 

Bag of Books	\CMTags{5 Uses, 10 coins, 2 Weight} 

 

When your bag of books contains just the right book for the subject you're Spouting Lore on, consult the book, mark off a use, and take +1 to your roll.

 

Antitoxin	\CMTags{10 coins} 

 

When you drink antitoxin, you're cured of one poison affecting you.

 

Dungeon Rations	\CMTags{Ration, 5 Uses, 3 coins, 1 Weight} 

 

Not tasty, but not bad either.

 

Personal Feast	\CMTags{Ration, 1 Use, 10 coins, 1 Weight} 

 

Ostentatious to say the least.

 

Dwarven Hardtack	\CMTags{Requires Dwarf, Ration, 7 Uses, 3 coins, 1 Weight} 

 

Dwarves say it tastes like home. Everyone else says it tastes like home, on fire, in a hog farm.

 

Elven Bread	\CMTags{Ration, 7 Uses, 10 coins, 1 Weight} 

 

Only the greatest of elf-friends are treated to this rare delicacy.

 

Halfling Pipeleaf	\CMTags{5 Uses, 5 coins, 1 Weight} 

 

When you share halfling pipeleaf with someone, expend two uses and take +1 forward to Parley with them.

 
\section{Poisons}  \index{Poisons} \index{Poisons}
 

Oil of Tagit	\CMTags{Dangerous, Applied, 15 coins, 0 Weight} 

 

The target falls into a light sleep

 

Bloodweed	\CMTags{Dangerous, Touch, 12 coins, 0 Weight} 

 

The target deals -1d4 damage ongoing until cured

 

Goldenroot	\CMTags{Dangerous, Applied, 20 coins, 0 Weight} 

 

The target treats the next creature they see as a trusted ally, until proved otherwise

 

Serpent's Tears	\CMTags{Dangerous, Touch, 10 coins, 0 Weight} 

 

Anyone dealing damage against the target rolls twice and takes the better result.

 
\section{Services}  \index{Services} \index{Services}
 

A week's stay at a peasant inn	\CMTags{14-Charisma coins} 

 

A week's stay at a civilized inn	\CMTags{30-Charisma coins} 

 

A week's stay at the fanciest inn in town	\CMTags{43-Charisma coins} 

 

A week's unskilled mundane labor	\CMTags{10 coins} 

 

A month's pay for enlistment in an army	\CMTags{30 coins} 

 

A custom item from a blacksmith	\CMTags{Base Item + 50 coins} 

 

A night's "companionship"	\CMTags{20-Charisma coins} 

 

An evening of song and dance	\CMTags{18-Charisma coins} 

 

Escort for a day along a bandit-infested road	\CMTags{20 coins} 

 

Escort for a day along a monster-infested road	\CMTags{54 coins} 

 

A run-of-the-mill killing	\CMTags{5 coins} 

 

An assassination	\CMTags{120 coins} 

 

Healing from a Chirurgeon	\CMTags{5 coins} 

 

A month's prayers for the departed	\CMTags{1 coins} 

 

Repairs to a mundane item	\CMTags{25\% of the item's cost} 

 
\section{Meals}  \index{Meals} \index{Meals}
 

A hearty meal for one	\CMTags{1 coins} 

 

A poor meal for a family	\CMTags{1 coins} 

 

A feast	\CMTags{15 coins per person} 

 
\section{Transport}  \index{Transport} \index{Transport}
 

Cart and Donkey	\CMTags{50 coins, Load 20} 

 

This donkey is sworn to carry your burdens.

 

Horse	\CMTags{75 coins, Load 10} 

 

Warhorse	\CMTags{400 coins, Load 12} 

 

Wagon	\CMTags{150 coins, Load 40} 

 

Barge	\CMTags{50 coins, Load 15} 

 

River boat	\CMTags{150 coins, Load 20} 

 

Merchant Ship	\CMTags{5,000 coins, Load 200} 

 

War Ship	\CMTags{20,000 coins, Load 100} 

 

Passage on a safe route	\CMTags{1 coins} 

 

Passage on a tough route	\CMTags{10 coins} 

 

Passage on a dangerous route	\CMTags{100 coins} 

 
\section{Land and Buildings}  \index{Land and Buildings} \index{Land} \index{Buildings}
 

A hovel	\CMTags{20 coins} 

 

A cottage	\CMTags{500 coins} 

 

A house	\CMTags{2,500 coins} 

 

A mansion	\CMTags{50,000 coins} 

 

A keep	\CMTags{75,000 coins} 

 

A castle	\CMTags{250,000 coins} 

 

A grand castle	\CMTags{1,000,000 coins} 

 

A month's upkeep	\CMTags{1\% of the cost} 

 
\section{Bribes}  \index{Bribes} \index{Bribes}
 

A peasant dowery	\CMTags{20-Charisma coins} 

 

"Protection" for a small business	\CMTags{100-Charisma coins} 

 

A government bribe	\CMTags{50-Charisma coins} 

 

A compelling bribe	\CMTags{80-Charisma coins} 

 

An offer you can't refuse	\CMTags{500-Charisma coins} 

 
\section{Gifts and Finery}  \index{Gifts and Finery} \index{Gifts} \index{Finery}
 

A peasant gift	\CMTags{1 coins} 

 

A fine gift	\CMTags{55 coins} 

 

A noble gift	\CMTags{200 coins} 

 

A ring or cameo	\CMTags{75 coins} 

 

Finery	\CMTags{105 coins} 

 

A fine tapestry	\CMTags{350+ coins} 

 

A crown fit for a king	\CMTags{5,000 coins} 

 
\section{Hoards}  \index{Hoards} \index{Hoards}
 

A goblin's stash	\CMTags{2 coins} 

 

A lizardman's trinkets	\CMTags{5 coins} 

 

A priceless sword	\CMTags{80 coins} 

 

An orc warchief's tribute	\CMTags{250 coins} 

 

A dragon's mound of coins and gems	\CMTags{130,000 coins} 

 
