\chapter{The Fighter}
 \index{The Fighter} \index{Fighter}
 





It’s a thankless job—living day to day by your armor and the skill of your arm. To dive heedlessly into danger. They won’t be playing golden horns for the time you took that knife to the ribs for them in the bar in Bucksberg. No flock of angels to sing of the time you dragged them, still screaming, from the edge of the Pits of Madness, no.

 

Forget them.

 

You do this for the guts and the glory. The scream of battle and the hot hot blood of it. You are a beast of iron. Your friends may carry blades of forged steel but, Fighter, you are steel. While your traveling companions might moan about their wounds over a campfire in the wilderness, you bear your scars with pride.

 

You are the wall—let every danger smash itself to nothing on you. In the end, you’ll be the last one standing.



 
\section{Names}  \index{Names} \index{Names}
 



{\em Dwarf} : Ozruk, Surtur, Brunhilda, Annika, Janos, Greta, Dim, Rundrig, Jarl, Xotoq

   

{\em Elf} : Elohiir, Sharaseth, Hasrith, Shevaral, Cadeus, Eldar, Kithracet, Thelian

     

{\em Halfling} : Finnegan, Olive, Randolph, Bartleby, Aubrey, Baldwin, Becca

  

{\em Human} : Hawke, Rudiger, Gregor, Brianne, Walton, Castor, Shanna, Ajax, Hob



 
\section{Look}  \index{Look}
 



Choose one for each:

 

Hard Eyes, Dead Eyes, or Eager Eyes

 

Wild Hair, Shorn Hair, or Battered Helm

 

Calloused Skin, Tanned Skin, or Scarred Skin

 

Built Body, Lithe Body, or Ravaged Body



 
\section{Stats}  \index{Stats} \index{Stats}
 



Assign these scores to your stats:

 

17 (+2), 15 (+1), 13 (+1), 11 (+0), 9 (+0), 8 (-1)

 

You start with 10+Constitution HP.



 

Your base damage is d10.

 
\section{Starting Moves}  \index{Starting Moves} \index{Moves}
 


\startInstructionsAfterHeader
Choose a racial move:
\stopInstructionsAfterHeader
 


\subsection{Dwarf}  \index{Dwarf} \index{Dwarf}
 

When you share a drink with someone, you may Parley with them with Con instead of Cha.

 
\subsection{Elf}  \index{Elf} \index{Elf}
 

Choose one weapon—you can always treat weapons of that type as if they had the Precise tag.

 
\subsection{Halfling}  \index{Halfling}
 

When you Defy Danger and use your small size to your advantage, take +1.

 
\subsection{Human}  \index{Human} \index{Human}
 

Once per battle you may choose a damage roll (yours or someone else's) and reroll it.



 


\startInstructions
You start with these moves:
\stopInstructions
 
\subsection{Bend Bars, Lift Gates (Str)}  \index{Bend Bars Lift Gates (Str)} \index{Bend} \index{Bars} \index{Lift} \index{Gates} \index{(str)}
 

When you use pure strength to destroy an inanimate obstacle, roll+Str. On a 10+ choose 3. On a 7-9 choose 2.

 
\startitemize[1,packed]

\item It doesn’t take a very long time

 
\item Nothing of value is damaged

 
\item It doesn’t make an inordinate amount of noise

 
\item You can fix the thing again without a lot of effort


\stopitemize
 
\subsection{Armored}  \index{Armored} \index{Armored}
 

You ignore the Clumsy tag on armor you wear.

 
\subsection{Signature Weapon}  \index{Signature Weapon} \index{Signature} \index{Weapon}
 

This is your weapon. There are many like it, but this one is yours. Your weapon is your best friend. It is your life. You master it as you master your life. Your weapon, without you, is useless. Without your weapon, you are useless. You must wield your weapon true.

 

Choose a base description, all are 2 weight:

 
\startitemize[1,packed]

\item Sword

 
\item Axe

 
\item Hammer

 
\item Spear

 
\item Flail

 
\item Fists


\stopitemize
 

Choose the range that best fits your weapon:

 
\startitemize[1,packed]

\item Hand

 
\item Close

 
\item Reach


\stopitemize
 

Choose two enhancements:

 
\startitemize[1,packed]

\item Hooks and spikes. +1 damage, but +1 weight.

 
\item Sharp. +2 piercing.

 
\item Perfectly weighted. +precise.

 
\item Serrated edges. +1 damage.

 
\item Glows in the presence of one type of creature, your choice.

 
\item Huge. +messy, +forceful.

 
\item Versatile. Choose another range.

 
\item Well-crafted. -1 weight.


\stopitemize
 

Choose a look:

 
\startitemize[1,packed]

\item Ancient

 
\item Unblemished

 
\item Ornate

 
\item Blood-stained

 
\item Sinister


\stopitemize




 
\section{Alignment}  \index{Alignment} \index{Alignment}
 
\startInstructionsAfterHeader
Choose an alignment:
\stopInstructionsAfterHeader
 


\subsection{Good}  \index{Good} \index{Good}
 

Defend those weaker than you

 
\subsection{Neutral}  \index{Neutral} \index{Neutral}
 

Defeat a worthy opponent

 
\subsection{Evil}  \index{Evil} \index{Evil}
 

Kill a defenseless or surrendered enemy



 
\section{Gear}  \index{Gear} \index{Gear}
 



Your Load is 9+Str. You carry your signature weapon and dungeon rations (1 weight, 5 uses). Choose your defenses:

 
\startitemize[1,packed]

\item Chainmail (1 armor, 1 weight) and Adventuring gear (1 weight)

 
\item Scale armor (2 armor, 3 weight)


\stopitemize
 

Choose two:

 
\startitemize[1,packed]

\item 2 Healing potions (2 weight)

 
\item shield (+1 armor, 2 weight)

 
\item Antitoxin, dungeon rations (1 weight), and poultices and herbs (1 weight)

 
\item 22 Gold


\stopitemize


 
\section{Bonds}  \index{Bonds} \index{Bonds}
 



Fill in the name of one of your companions in at least one:

 

\thinrules[2] owes me their life, whether they admit it or not.

 

I have sworn to protect \thinrules[2].

 

I worry about the ability of \thinrules[2] to survive in the dungeon.

 

\thinrules[2] is soft, but I will make them hard like me.



 
\section{Advanced Moves}  \index{Advanced Moves} \index{Advanced} \index{Moves}
 


\startInstructionsAfterHeader
When you gain a level from 2-5, choose from these moves.
\stopInstructionsAfterHeader
 
\subsection{Merciless}  \index{Merciless} \index{Merciless}
 

When you deal damage, deal +1d4 damage.

 
\subsection{Heirloom}  \index{Heirloom} \index{Heirloom}
 

When you consult the spirits of that reside within your signature weapon, roll+Cha. The spirits will give you an insight relating to the current situation, and might ask you some questions in return. On a 10+, the GM will give you good detail. On a 7-9, the GM will give you an impression.

 
\subsection{Armor Mastery}  \index{Armor Mastery} \index{Armor} \index{Mastery}
 

When you take damage you can choose to let your armor take the brunt of it. The damage is negated but your armor or shield (your choice) is -1 armor until you get it repaired at a smithy or workshop.

 
\subsection{Improved Weapon}  \index{Improved Weapon} \index{Improved} \index{Weapon}
 

Choose one extra enhancement for your signature weapon.

 
\subsection{Seeing Red}  \index{Seeing Red} \index{Red}
 

When you Discern Realities during combat, you take +1.

 
\subsection{Interrogator}  \index{Interrogator} \index{Interrogator}
 

When you Parley using threats of impending violence as leverage you may use Str instead of Cha.

 
\subsection{Scent of Blood}  \index{Scent of Blood} \index{Scent} \index{Blood}
 

When you Hack and Slash an enemy, you take +1d4 damage forward against that enemy.

 
\subsection{Multiclass Dabbler}  \index{Multiclass Dabbler} \index{Multiclass} \index{Dabbler}
 

Get one move from another class. For the purposes of Multiclass Dabbler the Wizard's Spellbook, Prepare Spells, and Cast a Spell moves count as one move. The Cleric's Commune and Cast a Spell moves also count as one move. If you gain the ability to cast spells you cast them as if you were one level lower.

 
\subsection{Iron Hide}  \index{Iron Hide} \index{Iron} \index{Hide}
 

You gain +1 armor.

 
\subsection{Blacksmith}  \index{Blacksmith} \index{Blacksmith}
 

When you have access to a forge you can graft the magical powers of a weapon onto your signature weapon. This process destroys the magical weapon. Your signature weapon gains the magical powers of the destroyed weapon.

 
\startInstructions
When you gain a level from 6-10, choose from these moves or the level 2-5 moves.
\stopInstructions
 
\subsection{Bloodthirsty}  \index{Bloodthirsty} \index{Bloodthirsty}
 

Replaces: Merciless

 

When you deal damage, deal +1d8 damage.

 
\subsection{Armored Perfection}  \index{Armored Perfection} \index{Armored} \index{Perfection}
 

Replaces: Armor Mastery

 

When you take damage you can choose to let your armor take the brunt of it. The damage is negated and you take +1 forward against the attacker but your armor or shield (your choice) is -1 armor until you get it repaired at a smithy or workshop.

 
\subsection{Evil Eye}  \index{Evil Eye} \index{Evil} \index{Eye}
 

Requires: Seeing Red

 

When you go into a dangerous situation, roll+Cha. On a 10+, hold 2. On a 7-9, hold 1. Spend your hold to make eye contact with an NPC present, who freezes or flinches and can't act until you break it off. On a miss, your enemies immediately identify you as their biggest threat.

 
\subsection{Taste of Blood}  \index{Taste of Blood} \index{Taste} \index{Blood}
 

Replaces: Scent of Blood

 

When you Hack and Slash an enemy, you take +1d8 damage forward against that enemy.

 
\subsection{Multiclass Initiate}  \index{Multiclass Initiate} \index{Multiclass} \index{Initiate}
 

Required: Multiclass Dabbler

 

Get one move from another class. Treat your level as one lower for choosing the move.

 
\subsection{Steel Hide}  \index{Steel Hide} \index{Steel} \index{Hide}
 

Replaces: Iron Hide

 

You gain +2 armor.

 
\subsection{Through Death's Eyes}  \index{Through Death's Eyes} \index{Death's} \index{Eyes}
 

When you go into battle, roll+Wis. On a 10+, name someone who will live and someone who will die. On a 7-9, name someone who will live or someone who will die. Name NPCs, not player characters. The GM will make your vision come true, if it's even remotely possible. On a 6- you see your own death and consequently take a -1 throughout the battle.

 
\subsection{Eye for Weaponry}  \index{Eye for Weaponry} \index{Eye} \index{Weaponry}
 

When you look over an enemy's weaponry, ask the GM how much damage they do.

 
\subsection{Superior Warrior}  \index{Superior Warrior} \index{Superior} \index{Warrior}
 

When you Hack and Slash on a 12+ you deal your damage, avoid their attack, and impress, dismay, or frighten your enemy.









 
